---
title: "MARC-35 Serum metabolome by RSV vs RV"
author: "Kohei Hasegawa"
date: "9/11/2018"
output: html_document
---

  **Aim(s):** 1) To examine the between-virus differences in metabolomie profiles; 2) their biologic interpretation. Prediction of RSV vs. RV is not the goal.
  
  **Patients:** 116 patients (excluding RSV-RV coinfection)
  
  **Exposure:** RSV-only vs. RV-only
  
  **Outcomes:** Serum metabolites (708)
  
  **Confounders:** Age, sex. 
  
  **Analysis:** 1) linear regression (after log-transformation) adjsuting for cnofounder(s) with FDR correction, 2) pathway analysis?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Upload packages
```{r warning=FALSE}
library("tidyverse")
library("readstata13")
library("broom")
```


## 2. Upload datasets
```{r}
# Metabolome data (all metabolites)
df_metabolome <- read_csv("../data/merged_serum_VirusCombined.csv") 
dim(df_metabolome) #; colnames(df_metabolome)

# Metadata
df_metadata <- read.dta13("../data/m35_metadata_n1016_2018march.dta")
dim(df_metadata)

# Merge the datasets
df_merged <- df_metabolome %>% inner_join(., df_metadata, by="study_id")
dim(df_merged)  # n=116

```

## 3. Data Tranformation
```{r}
# Log2 transformation of metabolome data, then replace the original data with transformed
# The metabolites are columns 3-710
df_merged[, 3:710] <- log(df_merged[3:710], base=2)
head(df_merged)
```

## 4. Prep for regression
```{r eval=FALSE}
# Set refelence level -- unlikely to be needed
#df_merged$Virus_Combined <- as_factor(df_merged$Virus_Combined)  # Change to factor
#df_merged$Virus_Combined <- relevel(df_merged$Virus_Combined, "Rvonly") 

```

## 5. Unadjusted regression
```{r}
# Develop a function to run UNadjusted model
lm_unadj_function <- function(y) lm(y ~ Virus_Combined, 
                                    data = df_merged)
# Apply the function to each column
#lapply (df_merged[,3:710], lm_unadj_function)

# Apply the function to the metabolites, and get FDR 
df_output_unadjusted <- map_dfr(df_merged[,3:710], 
                                function(x) summary(lm_unadj_function(x)) %>% 
          broom::tidy(conf.int=TRUE)) %>%
          filter(term == "Virus_CombinedRvonly")%>%   # Get outputs for the virus only (not intercepts or covariates)
          select(-c(term, std.error, statistic)) %>%  # Remove unnecessary variables
          mutate(metabolite <- colnames(df_merged[3:710])) %>% # Add names of metabolites
          mutate(fdr_p = p.adjust(p.value, "fdr")) %>% # Compute FDR
          filter(fdr_p < 0.05)   # Get only FDR <0.05

dim(df_output_unadjusted)  # 122 metabolites are significant
write_csv(df_output_unadjusted, "./output_unadjusted_fdr.csv")
```
```{r}
colnames(df_metadata)
summary(df_metadata$corticosteroids_pre_, useNA="ifany")
+ premature37 + corticosteroids_pre #+ CPAPintubate
```


## 6. Adjusted regression 
```{r}
# Develop a function to run Adjusted model
lm_adj_function <- function(y) lm(y ~ Virus_Combined
                                    + Age_mo + intake_sex
                                    + premature37 + corticosteroids_life #+ CPAPintubate
                                    , data = df_merged)

# Apply the function to the metabolites, and get FDR 
df_output_adjusted <- map_dfr(df_merged[,3:710], 
                                function(y) summary( lm_adj_function(y) ) %>% 
          broom::tidy(conf.int=TRUE)) %>%
          filter(term == "Virus_CombinedRvonly")%>%   # Get outputs for the virus only (not intercepts or covariates)
          select(-c(term, std.error, statistic)) %>%  # Remove unnecessary variables
          mutate(metabolite <- colnames(df_merged[3:710])) %>% # Add names of metabolites
          mutate(fdr_p = p.adjust(p.value, "fdr")) %>% # Compute FDR
          filter(fdr_p < 0.05)   # Get only FDR <0.05

dim(df_output_adjusted)  # 106 metabolites are significant
write_csv(df_output_adjusted, "output_adjusted_fdr.csv")
```


## Random-effect
```{r}
library(lme4)
library(lmerTest)
df_merged$site <- as.factor(df_merged$site)
df_merged$raceethn <- as.factor(df_merged$raceethn)

# place holder
coef_lm    <- 708
p.value_lm <- 708
# For Loop
for (i in names(df_merged)[3:710]) {
      fit <- lmer(df_merged[[i]] ~ Virus_Combined
                            + Age_mo + intake_sex 
                            + raceethn
                            + premature37 + corticosteroids_life  #+ CPAPintubate
                            + (1 | site), 
                            data = df_merged, REML = FALSE)
       coef_lm[i]    <- summary(fit)$coefficients[2,1]  # Summary of 2nd row + 1st column = coef 
       p.value_lm[i] <- summary(fit)$coefficients[2,5]  # Summary of 2nd row + 5th column = p-value 
    }


df_random_output_orig <- cbind(coef_lm, p.value_lm) %>% tbl_df()
df_random_output <- df_random_output_orig[2:709,] %>%
          mutate(metabolite <- colnames(df_merged[3:710])) %>% # Add names of metabolites
          mutate(fdr_p = p.adjust(p.value_lm, "fdr")) %>% # Compute FDR
          filter(fdr_p < 0.05)   # Get only FDR <0.05
dim(df_random_output) # 152 metabolites are significant
write_csv(df_random_output, "output_random_adjusted_fdr.csv")
```





