---
title: "MARC-35 Serum metabolome by RSV vs RV"
author: "Kohei Hasegawa"
date: "9/11/2018"
output: html_document
---

  **Aim(s):** 1) To examine the between-virus differences in metabolomie profiles; 2) their biologic interpretation. Prediction of RSV vs. RV is not the goal.
  **Patients:** 116 patients (excluding RSV-RV coinfection)
  **Exposure:** RSV-only vs. RV-only
  **Outcomes:** Serum metabolites (708)
  **Confounders:** Age, sex. 
  **Analysis:** 1) linear regression (after log-transformation) adjsuting for cnofounder(s) with FDR correction, 2) pathway analysis?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Upload packages
```{r}
library("tidyverse")
library("readstata13")
library("broom")
```


## 2. Upload datasets
```{r}
# Metabolome data (all metabolites)
df_metabolome <- read_csv("../data/merged_serum_VirusCombined.csv") 
dim(df_metabolome) #; colnames(df_metabolome)

# Metadata
df_metadata <- read.dta13("../data/m35_metadata_n1016_2018march.dta")
dim(df_metadata)

# Merge the datasets
df_merged <- df_metabolome %>% inner_join(., df_metadata, by="study_id")
dim(df_merged)  # n=116

```

## 3. Data Tranformation
```{r}
# Log2 transformation of metabolome data, then replace the original data with transformed
# The metabolites are columns 3-710
df_merged[, 3:710] <- log(df_merged[3:710], base=2)
head(df_merged)
```

## 4. Prep for regression
```{r}
# Set refelence level
#df_merged$Virus_Combined <- as_factor(df_merged$Virus_Combined)  # Change to factor
#df_merged$Virus_Combined <- relevel(df_merged$Virus_Combined, "Rvonly") 

```

## 5. Unadjusted regression
```{r TRIAL -- to be deleted}
fit3<- lm(df_merged[[3]] ~Virus_Combined,
         data=df_merged)
tidy(fit3)
summary(fit3)$coefficients[2,4]

fit4<- lm(df_merged[[4]] ~Virus_Combined,
         data=df_merged)
tidy(fit4)
summary(fit4)$coefficients[2,4]
```
```{r}
# Develop a function to run unadjusted model
lm_function <- function(y) lm(y ~ Virus_Combined, 
                           data = df_merged)
# Apply the function to each column
lapply (df_merged[,3:6], lm_function)
map_dfr(df_merged[,3:6], function(x) summary(lm_function(x)) %>% glance()) 
df_ouput_unadjusted <- map_dfr(df_merged[,3:6], function(x) summary(lm_function(x)) %>% 
          tidy()) %>% 
          filter(term == "Virus_CombinedRvonly")

rownames(ouput_unadjusted) <-colnames(df_merged[3:6])
```



```{r}
number_metabolites <- 2

# placeholders
coef_lm    <- numeric(number_metabolites)
se_lm      <- numeric(number_metabolites)
p.value_lm <- numeric(number_metabolites)

View(head(df_merged))

# for loop
for (i in seq_along(colnames(df_merged)[3:4])){
    fit <- lm(df_merged[[i]] ~ Virus_Combined,
              data=df_merged)  
    coef_lm[i]    <- summary(fit)$coefficients[2,1]  # Summary of 2nd row + 1st column = coef 
    se_lm[i]      <- summary(fit)$coefficients[2,2]  # Summary of 2nd row + 1st column = se 
    p.value_lm[i] <- summary(fit)$coefficients[2,4]  # Summary of 2nd row + 1st column = p-value 
}
    
    
output_lm <- cbind(colnames(df_merged)[3:4], 
                   coef_lm, se_lm, p.value_lm) %>% tbl_df()
output_lm$coef_lm    <- as.numeric(output_lm$coef_lm) %>% round (., digits =2)
output_lm$se_lm      <- as.numeric(output_lm$se_lm)   %>% round (., digits =2)
output_lm$p.value_lm <- as.numeric(output_lm$p.value_lm)
output_lm$fdr_p <- as.numeric(output_lm$p.value_lm)   %>% p.adjust(., "fdr") %>% round(., digits=4)
output_lm$p.value_lm <- output_lm$p.value_lm %>% round(., digits=4)
output_lm; 

  output_lm %>% filter(., fdr_p <0.05) 
    
```


